---
title: "Covid91 vaccine study Final2022"
author: "Trevor Ragazzo"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```



```{r}
FinalDat = read.csv(file="final_Data.csv")
```


```{r}
Males = filter(FinalDat,  sex == "M")
```

```{r}
Males = select(Males, -sex)
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
dd2 <- FinalDat %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
 In this first graph we see it showing us that half of the participants took the covid 19 vacine and the other half took a placebo. With this information we get a result of 70% of the people that took the placebo got covid while 30% people with the vaccine got covid. We saw a greater result of people with covid that did not get the vaccine. 
  
  
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

This numerical table shows us a result of the people who took the placebo and got covid was 2.63% and the people with the vaccine was about 1.28%. Which is showing the vaccine to be effective. 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



## Females


```{r}
Females = filter(FinalDat,  sex == "F")
```

```{r}
Females = select(Females, -sex)
```

```{r}
dd2 <- FinalDat %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```



## LGBTQ



## Druggies

