---
title: "Covid91 vaccine study Final2022"
author: "Trevor Ragazzo"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```







# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
Males = filter(FinalDat,  sex == "M")
```

```{r}
Males = select(Males, -sex)
```

## Males Graphical Results

```{r}
dd2 <- FinalDat %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
 In this first graph we see it showing us that half of the participants took the covid 19 vacine and the other half took a placebo. With this information we get a result of 70% of the people that took the placebo got covid while 30% people with the vaccine got covid. We saw a greater result of people with covid that did not get the vaccine. 
  
  
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

This numerical table shows us a result of the people who took the placebo and got covid was 2.63% and the people with the vaccine was about 1.28%. Which is showing the vaccine to be effective. 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fishtestm=fisher.test(table1)
print(fishtestm)
```

The fisher exact test is showing us that the participants that took the placebo is almost 2 times more likely to catch covid rather than the people who had the vaccine. With such a low P-value it means we cannot recreate this experiment with the vaccine and placebo being the same, and getting the results we have just observed. 



```{r}
options(digits=1)
maleffic=fishtestm$estimate/(1 + fishtestm$estimate)*100
malefficlo=fishtestm$conf.int[1]/(1+fishtestm$conf.int[1])*100
maleffichi=fishtestm$conf.int[2]/(1+fishtestm$conf.int[2])*100
print(maleffic)
print(malefficlo)
print(maleffichi)
```

we see the confidence interval to be `r maleffichi` and the lower to be `r malefficlo`



## Females


```{r}
Females = filter(FinalDat,  sex == "F")
```

```{r}
Females = select(Females, -sex)
```

```{r}
dd2 <- FinalDat %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart is showing us results of females that took a covid vaccine and those who took a placebo. With this information we get results of 60% of the people that took the placebo got covid, while the people with the vaccine only 40% got covid. 

  

```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```

From the numerical results we see that around 2.3% of the people who received the placebo actually got covid. While the people who received the covid vaccine only 1.62% got covid. We see that the vaccine seem to be working, almost 2 times better than the placebo.


```{r}
chisq.test(table2)
chisqtestGC(table2)
fishtestf=fisher.test(table2)
fishtestf
```

From this test we see that if you have taken the placebo you are 1.42 times more likely to get covid compare to the people that took the covid vaccine. We also notice a low P-Value, this means we would not gte the same results in this test if the placebo and vaccine were the same


```{r}
options(digits=1)
femeffic=fishtestf$estimate/(1 + fishtestf$estimate)*100
femefficlo=fishtestf$conf.int[1]/(1+fishtestf$conf.int[1])*100
femeffichi=fishtestf$conf.int[2]/(1+fishtestf$conf.int[2])*100
print(femeffic)
print(femefficlo)
print(femeffichi)
```




## LGBTQ



```{r}
LGBTQ = filter(FinalData, LGBTQ == "gay")
```

```{r}
dd2 <- LGBTQ %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

From this bar chart we see results of the members of the LGBTQ that received the vaccine actually got higher rates of covid. 85% of LGBTQ that took the vaccine got covid, seems like placebo might of had a better effect. 

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```

From the numerical summaries we see that about 3% of LGBTQ that had the vaccine got covid, while the ones who took the placebo was only about .05%. 

```{r}
chisq.test(table3)
chisqtestGC(table3)
fishtestL=fisher.test(table3)
fishtestL
```

We see the odd ration to be .1 this tells us that people who take the placebo has a .1% of catching covid. 

```{r}
options(digits=1)
Lemeffic=fishtestL$estimate/(1 + fishtestL$estimate)*100
Lemefficlo=fishtestL$conf.int[1]/(1+fishtestL$conf.int[1])*100
Lemeffichi=fishtestL$conf.int[2]/(1+fishtestL$conf.int[2])*100
print(Lemeffic)
print(Lemefficlo)
print(Lemeffichi)
```

This efficiency ratio of 13 means the vaccine was only 13% efficient. The confidence ratio for the vaccine shows to be `r Lemefficlo` and the high end will be `r Lemeffichi`



## Druggies


```{r}
DrugUser = filter(FinalData, DrugUser == "yes")
```

## DrugUser Graphical Description

```{r}
dd2 <- DrugUser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart is showing results that the drugusers that took the vaccine was more likely to get covid rather than taking the placebo. About 85% of drug users that took the vaccine got covid and the people that ook the placebo was only 15% likely. 

## DrugUser Numerical Results

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table4)
colPerc(table4)
```

From the numerical results we see that 3% of drugusers that took the vaccin got covid. While only .6 % of drugusers that took the placebo got covid. From these results we see that with the vaccine you are more than 3 times more likely to get covid. 



## DrugUser Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fishtestD=fisher.test(table4)
fishtestD
```

We see from the Inferential Results that the drugusers that took the placebo were about .2 times less likely to get covid rather than the drugusers that took the vaccine. With this low P-Value we can determine that the placebo and the vaccine are indeed different and we couldnt get these results if they were the same. 

## DrugUser Efficiency

```{r}
options(digits=1)
Deffic=fishtestD$estimate/(1 + fishtestD$estimate)*100
Defficlo=fishtestD$conf.int[1]/(1+fishtestD$conf.int[1])*100
Deffichi=fishtestD$conf.int[2]/(1+fishtestD$conf.int[2])*100
print(Deffic)
print(Defficlo)
print(Deffichi)
```

From the Efficiency we see that the vaccine is 15% effective. We see a confidence interval high of `r Deffichi` and a low of `r Defficlo`






